---
title: "Sky/Loch SR/ST collation and metadata for Aquatic Surface Ref Request"
author: "B Steele"
date: "2023-03-03"
output: html_document
---

# Purpose

To process the C2 SR/ST pull and get a scene list for Sky/Loch for the Provisional Aquatic SR product.

# Setup

```{r}
library(googledrive)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(gghighlight)

data_dir = file.path('data/remoteSensing/')
drive_dir = file.path('data/fromDrive')
drive_auth()
1
```

# Download and collate data and metadata from Drive

Download and collate data and metadata into separate files.

```{r}
#get a file list
files = drive_ls(path = 'sky_loch')

#function for downloading to data folder
dr_down = function(filename, fileid){
  drive_download(file = as_id(fileid), path = file.path(drive_dir, filename), overwrite = T)
}

#map over the function to download all files
map2(files$name, files$id, dr_down)

# create a list of the files in the tmp directory
list = list.files(drive_dir)
#add prefix
list = file.path(drive_dir, list)

meta_list = list[grepl('meta', list) & !grepl('LEVEL1', list)]
data_list = list[!grepl('meta', list)]
L1_list = list[grepl('LEVEL1', list)]

#read them in and map to a dataframe
collated_data = map_dfr(data_list, read_csv) %>% 
  filter(!is.na(med_Blue))
collated_metadata = map_dfr(meta_list, read_csv)
L1_list = map_dfr(L1_list, read_csv)

#clean up workspace
rm(files)
```

Reformat the data system:index so that it will play nicely with the metadata and so we pull out the site rowid.

```{r}
grabRowid = function(sys_idx){
  parsed = str_split(sys_idx, '_')
  str_len = length(unlist(parsed))
  unlist(parsed)[str_len]
}

grabSystemIndex = function(sys_idx){
  parsed = str_split(sys_idx, '_')
  str_len = length(unlist(parsed))
  parsed_sub = unlist(parsed)[1:(str_len-1)]
  str_flatten(parsed_sub, collapse = '_')
}

collated_data$rowid = map(collated_data$`system:index`,grabRowid)
collated_data$`system:index` = map(collated_data$`system:index`, grabSystemIndex)

collated_data$`system:index` = as.character(collated_data$`system:index`)

```

Grab only the metadata we want

```{r}
filtered_metadata <- collated_metadata %>% 
  mutate(IMAGE_QUALITY = if_else(is.na(IMAGE_QUALITY), IMAGE_QUALITY_OLI, IMAGE_QUALITY)) %>% 
  select(`system:index`, 
         LANDSAT_PRODUCT_ID,
         WRS_PATH, 
         WRS_ROW, 
         'mission' = SPACECRAFT_ID, 
         'date' = DATE_ACQUIRED, 
         'UTC_time' = SCENE_CENTER_TIME, 
         CLOUD_COVER,
         IMAGE_QUALITY, 
         IMAGE_QUALITY_TIRS, 
         SUN_AZIMUTH, 
         SUN_ELEVATION) 
filtered_L1 <- L1_list %>% 
  rename(IMAGE_QUALITY = IMAGE_QUALITY_OLI) %>% 
  select(`system:index`, 
         LANDSAT_PRODUCT_ID,
         WRS_PATH, 
         WRS_ROW, 
         'mission' = SPACECRAFT_ID, 
         'date' = DATE_ACQUIRED, 
         'UTC_time' = SCENE_CENTER_TIME, 
         CLOUD_COVER,
         IMAGE_QUALITY, 
         IMAGE_QUALITY_TIRS, 
         SUN_AZIMUTH, 
         SUN_ELEVATION) 
  
```

Join the data and metadata.

```{r}
data = left_join(collated_data, filtered_metadata) %>% 
  mutate(rowid = as.character(rowid))

#clean up workspace
rm(collated_data, collated_metadata)

write.csv(data, file.path(data_dir, paste0('SkyLoch_LandsatC2_SRST_collated_v', Sys.Date(), '.csv')))
```

## Get unique system IDs for Provisional Aquatic SR product

```{r}
filtered_forL1 = filtered %>% 
  select(-LANDSAT_PRODUCT_ID)
L1_names = filtered_L1 %>% 
select(date, LANDSAT_PRODUCT_ID, WRS_PATH, WRS_ROW) %>% 
  left_join(filtered_forL1, .)
uindex = unique(L1_names$LANDSAT_PRODUCT_ID)
```

Save as a txt file

```{r}
uindex = tibble(uindex)
write_delim(uindex, col_names = F, file.path(data_dir, 'SkyLoch_L1_SceneList_forProvisional.txt'))
```
