---
title: "Application of Handoff Coefficients"
author: "B Steele"
date: "2023-06-20"
output: html_document
---


```{r}
library(tidyverse)

#point to directories
rs_dir = 'data/remoteSensing/'
```

# Purpose

This script applies the regional hand-off coefficients to the Sky/Loch LS stack.

## Load files

The handoff coefficients here were calculated using all lakes greater than or equal to 1 ha in EcoRegion 3, Zone 21 following the methods described in Topp, et al. (2020).

```{r}
handoff = read.csv(file.path(rs_dir, 'ER3_Z21_regional_handoff_coefficients_v2023-05-24.csv')) %>% 
  rename(mission = SatCorr)

point_rs = read.csv(file.path(rs_dir, 'SkyLoch_LandsatC2_SRST_collated_v2023-06-20.csv')) 

```

Because we're really only applying corrections to the median band values, let's drop all the others and pull the date/loc/mission to the front and drop Landsat 4.

```{r}
point_rs_sub = point_rs %>% 
  select(date, rowid,
         mission,
         med_Blue:med_Swir2,
         min_SurfaceTemp, med_SurfaceTemp, max_SurfaceTemp,
         prop_clouds:system.index,
         CLOUD_COVER:SUN_ELEVATION) %>% 
  filter(mission != 'LANDSAT_4')
```

## Apply coefficients

We'll use the regional coefficients to calculate Rrs values for LS 5, 8, 9 as relative values to LS 7.

Let's re-orient the handoff calcs so that we can just join with the upstream dataset and apply in a single step.

```{r}
handoff_h = handoff %>% 
  pivot_longer(names_to = 'coeff',
               values_to = 'value',
               cols = c('intercept', 'B1', 'B2')) %>% 
  pivot_wider(names_from = c('band', 'coeff'),
              values_from = 'value')
```

Now join with the rs subset

```{r}
point_rs_sub = full_join(point_rs_sub, handoff_h)
```

And apply the coefficients, and fill in LS 7 as corrected (all values are corrected *to* LS 7 values)

```{r}
point_rs_sub = point_rs_sub %>%
  mutate(med_Blue_corr = med_Blue_intercept + med_Blue_B1*med_Blue + med_Blue_B2*med_Blue^2,
         med_Red_corr = med_Red_intercept + med_Red_B1*med_Red + med_Red_B2*med_Red^2,
         med_Green_corr = med_Green_intercept + med_Green_B1*med_Green + med_Green_B2*med_Green^2,
         med_Nir_corr = med_Nir_intercept + med_Nir_B1*med_Nir + med_Nir_B2*med_Nir^2,
         med_Swir1_corr = med_Swir1_intercept + med_Swir1_B1*med_Swir1 + med_Swir1_B2*med_Swir1^2,
         med_Swir2_corr = med_Swir2_intercept + med_Swir2_B1*med_Swir2 + med_Swir2_B2*med_Swir2^2) %>%
  mutate(med_Blue_corr = ifelse(mission == 'LANDSAT_7', med_Blue, med_Blue_corr),
         med_Red_corr = ifelse(mission == 'LANDSAT_7', med_Red, med_Red_corr),
         med_Green_corr = ifelse(mission == 'LANDSAT_7', med_Green, med_Green_corr),
         med_Nir_corr = ifelse(mission == 'LANDSAT_7', med_Nir, med_Nir_corr),
         med_Swir1_corr = ifelse(mission == 'LANDSAT_7', med_Swir1, med_Swir1_corr),
         med_Swir2_corr = ifelse(mission == 'LANDSAT_7', med_Swir2, med_Swir2_corr))
```

## Filter scene summaries

Filter:

-   each scene-location must have at least 6 pixels in pCount_dswe1 (confident water) contributing to values

-   each scene must have an image quality of 7 or greater

-   data mush be in feasible range (0 to 0.20)


```{r}
filtered_corrected = point_rs_sub %>% 
  filter(!is.na(med_Blue_corr), pCount_dswe1 >= 6, IMAGE_QUALITY >= 7) %>% 
  filter_at(vars(med_Red_corr, med_Green_corr, med_Blue_corr, med_Nir_corr, med_Swir1_corr, med_Swir2_corr), all_vars(.<0.2 & . > 0.0))
```

Read in location info

```{r}
locs = read.csv(file.path('data', 'location_lat_longs_sky_loch.csv')) %>% 
  rowid_to_column() %>% 
  select(rowid, location)
```

Join filtered data with location info

```{r}
filtered_corrected <- filtered_corrected %>% 
  mutate(rowid = as.integer(rowid),
         sat = as.character(mission)) %>% 
  full_join(., locs)
```

## Export file

```{r}
filtered_corrected %>%
  select(date, rowid, location,
         mission,
         med_Blue_corr:med_Swir2_corr,
         min_SurfaceTemp:max_SurfaceTemp,
         prop_clouds:SUN_ELEVATION) %>%
  write.csv(., file.path(rs_dir, paste0('SkyLoch_LandsatC2_SRST_filtered_gt0_corr_', Sys.Date(), '.csv')), row.names = F)

```
