---
title: "Calculation of dominant wavelength"
author: "B Steele"
date: "2023-05-10"
output: html_document
---

```{r}
library(tidyverse)
library(colorscience)
library(ggthemes)
library(trend)

ls_dir = 'data/remoteSensing/'
```

# Purpose

This script calculates dominant wavelength (dWL) for Loch and Sky Ponds.

## Read in files

```{r}
files <- list.files(ls_dir)
filtered <- files[grepl('filtered_gt0', files)]

ls <- read.csv(file.path(ls_dir, filtered))
```

# Calculate dWL from adjusted values

Calculate Forel-Ule index - this is copied from <https://github.com/GlobalHydrologyLab/LakeReflectanceRepo/blob/master/00_Functions.R>

```{r}
fui.hue <- function(R, G, B) {
  
  # Convert R,G, and B spectral reflectance to dominant wavelength based
  # on CIE chromaticity color space
  
  # see Wang et al 2015. MODIS-Based Radiometric Color Extraction and
  # Classification of Inland Water With the Forel-Ule
  # Scale: A Case Study of Lake Taihu
  
  require(colorscience)
  # chromaticity.diagram.color.fill()
  Xi <- 2.7689*R + 1.7517*G + 1.1302*B
  Yi <- 1.0000*R + 4.5907*G + 0.0601*B
  Zi <- 0.0565*G + 5.5943*B
  
  # calculate coordinates on chromaticity diagram
  x <-  Xi / (Xi + Yi +  Zi)
  y <-  Yi / (Xi + Yi +  Zi)
  z <-  Zi / (Xi + Yi +  Zi)
  
  # calculate hue angle
  alpha <- atan2((x - 0.33), (y - 0.33)) * 180/pi
  
  # make look up table for hue angle to wavelength conversion
  cie <- cccie31 %>%
    mutate(a = atan2( (x - 0.33), (y - 0.33)) * 180/pi) %>%
    dplyr::filter(wlnm <= 700) %>%
    dplyr::filter(wlnm >=380)
  
  # find nearest dominant wavelength to hue angle
  wl <- cie[as.vector(sapply(alpha,function(x) which.min(abs(x - cie$a)))), 'wlnm']
  
  #out <- cbind(as.data.frame(alpha), as.data.frame(wl))
  
  return(wl)
}

## Code for connnecting dWL to forel-ule index 
fui.lookup <- tibble(dWL = c(471:583), fui = NA)

fui.lookup$fui[fui.lookup$dWL <= 583] = 21
fui.lookup$fui[fui.lookup$dWL <= 581] = 20
fui.lookup$fui[fui.lookup$dWL <= 579] = 19
fui.lookup$fui[fui.lookup$dWL <= 577] = 18
fui.lookup$fui[fui.lookup$dWL <= 575] = 17
fui.lookup$fui[fui.lookup$dWL <= 573] = 16
fui.lookup$fui[fui.lookup$dWL <= 571] = 15
fui.lookup$fui[fui.lookup$dWL <= 570] = 14
fui.lookup$fui[fui.lookup$dWL <= 569] = 13
fui.lookup$fui[fui.lookup$dWL <= 568] = 12
fui.lookup$fui[fui.lookup$dWL <= 567] = 11
fui.lookup$fui[fui.lookup$dWL <= 564] = 10
fui.lookup$fui[fui.lookup$dWL <= 559] = 9
fui.lookup$fui[fui.lookup$dWL <= 549] = 8
fui.lookup$fui[fui.lookup$dWL <= 530] = 7
fui.lookup$fui[fui.lookup$dWL <= 509] = 6
fui.lookup$fui[fui.lookup$dWL <= 495] = 5
fui.lookup$fui[fui.lookup$dWL <= 489] = 4
fui.lookup$fui[fui.lookup$dWL <= 485] = 3
fui.lookup$fui[fui.lookup$dWL <= 480] = 2
fui.lookup$fui[fui.lookup$dWL <= 475 & fui.lookup$dWL >470] = 1
```

## Apply dWL calculation

```{r}
ls_adj <- ls_adj %>% 
  mutate(dWL = fui.hue(Red_adj, Green_adj, Blue_adj))
```

# Quick visualization of dWL over time

```{r}
ls_gt0 = ls_adj %>% 
  mutate(date = as.Date(date),
         month = as.numeric(format(date, '%m'))) %>% 
  filter(med_SurfaceTemp > 283.15) 

ggplot(ls_gt0, aes(x = date, y = dWL, color = location)) +
  geom_point() +
  theme_bw() +
  #geom_smooth(method = 'lm', se = F) +
  scale_color_colorblind()
```

## Check to see if there is a trend in the The Loch

Let's just do a simple lm here, and adjust later if there does seem to be a trend.

```{r}
Loch = ls_gt0 %>% 
  filter(location == 'TheLoch') %>% 
  arrange(date)

dWL_Loch_lm = lm(Loch$dWL ~ Loch$date)
summary(dWL_Loch_lm)

Sky = ls_gt0 %>% 
  filter(location == 'SkyPond') %>% 
  arrange(date)

dWL_Sky_lm = lm(Sky$dWL ~ Sky$date)
summary(dWL_Sky_lm)
```

Does this change if we do an annual timeseries?

```{r}
Loch_annual = Loch %>% 
  mutate(year = as.numeric(format(date, '%Y'))) %>% 
  group_by(year) %>% 
  dplyr::summarize(med_dWL = median(dWL))
```

There are some gaps, so let's grab form 1992 forward (where there are no gaps)

```{r}
Loch_annual_sub = Loch_annual %>% filter(year >= 1992)

Loch_annual_ts = ts(Loch_annual_sub$med_dWL, start = 1992, end = 2022)

sens.slope(Loch_annual_ts, conf.level = 0.95)
plot(Loch_annual_ts)
```
