---
title: "Calculation of dominant wavelength"
author: "B Steele"
date: "2023-05-10"
output: html_document
---

```{r}
library(tidyverse)
library(colorscience)
library(ggthemes)
library(trend)

ls_dir = 'data/remoteSensing/'
```

# Purpose

This script calculates dominant wavelength (dWL) for Loch and Sky Ponds.

## Read in files

```{r}
files <- list.files(ls_dir)
filtered <- files[grepl('filtered', files)]

ls <- read.csv(file.path(ls_dir, filtered))
```

# Adjust Rrs values so no data are negative

```{r}
#get all unique dates
datelist <- unique(ls$date)

#look at range for all bands
range(ls$med_Blue)
range(ls$med_Green)
range(ls$med_Red)
range(ls$med_Nir)
range(ls$med_Swir1)
range(ls$med_Swir2)

```

## Make helper function to grab lowest value in any band

```{r}
getBandMinimum = function(oneDate) {
  mins = oneDate %>% 
    group_by(mission) %>% 
    dplyr::summarize(min = min(med_Blue, med_Green, med_Red, med_Nir, med_Swir1, med_Swir2))
  mins
}
```

## And then a function to adjust the band values to be positive

Because all values are well under .2, we're just going to add the minimum value to all bands, assuming the atmospheric correction bias effects all bands equally.

```{r}
adjustBands = function(dt) {
  df = ls %>% filter(date == dt)
  mins = getBandMinimum(df)
  df = df %>% 
    full_join(., mins) %>% 
    mutate(Red_adj = if_else(min < 0, med_Red + (-min), med_Red),
           Green_adj = if_else(min < 0, med_Green + (-min), med_Green),
           Blue_adj = if_else(min < 0, med_Blue + (-min), med_Blue),
           Nir_adj = if_else(min < 0, med_Nir + (-min), med_Nir),
           Swir1_adj = if_else(min < 0, med_Swir1 + (-min), med_Swir1),
           Swir2_adj = if_else(min < 0, med_Swir2 + (-min), med_Swir2))
  df           
}
```

## Apply adjustment across all dates

```{r}
ls_adj = map_dfr(datelist, adjustBands)
```

# Calculate dWL from adjusted values

Calculate Forel-Ule index - this is copied from <https://github.com/GlobalHydrologyLab/LakeReflectanceRepo/blob/master/00_Functions.R>

```{r}
fui.hue <- function(R, G, B) {
  
  # Convert R,G, and B spectral reflectance to dominant wavelength based
  # on CIE chromaticity color space
  
  # see Wang et al 2015. MODIS-Based Radiometric Color Extraction and
  # Classification of Inland Water With the Forel-Ule
  # Scale: A Case Study of Lake Taihu
  
  require(colorscience)
  # chromaticity.diagram.color.fill()
  Xi <- 2.7689*R + 1.7517*G + 1.1302*B
  Yi <- 1.0000*R + 4.5907*G + 0.0601*B
  Zi <- 0.0565*G + 5.5943*B
  
  # calculate coordinates on chromaticity diagram
  x <-  Xi / (Xi + Yi +  Zi)
  y <-  Yi / (Xi + Yi +  Zi)
  z <-  Zi / (Xi + Yi +  Zi)
  
  # calculate hue angle
  alpha <- atan2((x - 0.33), (y - 0.33)) * 180/pi
  
  # make look up table for hue angle to wavelength conversion
  cie <- cccie31 %>%
    mutate(a = atan2( (x - 0.33), (y - 0.33)) * 180/pi) %>%
    dplyr::filter(wlnm <= 700) %>%
    dplyr::filter(wlnm >=380)
  
  # find nearest dominant wavelength to hue angle
  wl <- cie[as.vector(sapply(alpha,function(x) which.min(abs(x - cie$a)))), 'wlnm']
  
  #out <- cbind(as.data.frame(alpha), as.data.frame(wl))
  
  return(wl)
}

## Code for connnecting dWL to forel-ule index 
fui.lookup <- tibble(dWL = c(471:583), fui = NA)

fui.lookup$fui[fui.lookup$dWL <= 583] = 21
fui.lookup$fui[fui.lookup$dWL <= 581] = 20
fui.lookup$fui[fui.lookup$dWL <= 579] = 19
fui.lookup$fui[fui.lookup$dWL <= 577] = 18
fui.lookup$fui[fui.lookup$dWL <= 575] = 17
fui.lookup$fui[fui.lookup$dWL <= 573] = 16
fui.lookup$fui[fui.lookup$dWL <= 571] = 15
fui.lookup$fui[fui.lookup$dWL <= 570] = 14
fui.lookup$fui[fui.lookup$dWL <= 569] = 13
fui.lookup$fui[fui.lookup$dWL <= 568] = 12
fui.lookup$fui[fui.lookup$dWL <= 567] = 11
fui.lookup$fui[fui.lookup$dWL <= 564] = 10
fui.lookup$fui[fui.lookup$dWL <= 559] = 9
fui.lookup$fui[fui.lookup$dWL <= 549] = 8
fui.lookup$fui[fui.lookup$dWL <= 530] = 7
fui.lookup$fui[fui.lookup$dWL <= 509] = 6
fui.lookup$fui[fui.lookup$dWL <= 495] = 5
fui.lookup$fui[fui.lookup$dWL <= 489] = 4
fui.lookup$fui[fui.lookup$dWL <= 485] = 3
fui.lookup$fui[fui.lookup$dWL <= 480] = 2
fui.lookup$fui[fui.lookup$dWL <= 475 & fui.lookup$dWL >470] = 1
```

## Apply dWL calculation

```{r}
ls_adj <- ls_adj %>% 
  mutate(dWL = fui.hue(Red_adj, Green_adj, Blue_adj))
```

# Quick visualization of dWL over time

```{r}
ls_gt0 = ls_adj %>% 
  mutate(date = as.Date(date),
         month = as.numeric(format(date, '%m'))) %>% 
  filter(med_SurfaceTemp > 283.15) 

ggplot(ls_gt0, aes(x = date, y = dWL, color = location)) +
  geom_point() +
  theme_bw() +
  #geom_smooth(method = 'lm', se = F) +
  scale_color_colorblind()
```

## Check to see if there is a trend in the The Loch

Let's just do a simple lm here, and adjust later if there does seem to be a trend.

```{r}
Loch = ls_gt0 %>% 
  filter(location == 'TheLoch') %>% 
  arrange(date)

dWL_Loch_lm = lm(Loch$dWL ~ Loch$date)
summary(dWL_Loch_lm)

Sky = ls_gt0 %>% 
  filter(location == 'SkyPond') %>% 
  arrange(date)

dWL_Sky_lm = lm(Sky$dWL ~ Sky$date)
summary(dWL_Sky_lm)
```

Does this change if we do an annual timeseries?

```{r}
Loch_annual = Loch %>% 
  mutate(year = as.numeric(format(date, '%Y'))) %>% 
  group_by(year) %>% 
  dplyr::summarize(med_dWL = median(dWL))
```

There are some gaps, so let's grab form 1992 forward (where there are no gaps)

```{r}
Loch_annual_sub = Loch_annual %>% filter(year >= 1992)

Loch_annual_ts = ts(Loch_annual_sub$med_dWL, start = 1992, end = 2022)

sens.slope(Loch_annual_ts, conf.level = 0.95)
plot(Loch_annual_ts)
```
